Cyclone: Fault Tolerance Middleware for NVM Clusters

### 2016-06-21 paper/paper.tex

In order to ease this bottleneck, Cyclone allows clients to assist in
replication. To do so with minimum changes to the RAFT protocol, the leader
continues to replicate entries to clients on a timeout, with the clients
providing best effort replication. The flow for assisted replication is
simple. For any RPC call added to the log, the leader sends back a message to
the client with a list of replicas. The client then sends the RPC call to
follower replicas on behalf of the leader. It then then waits for a timeout,
gathers and sends back a message to the leader with the list of replicas that
have responded within the timeout. Client assisted replication is therefore best
effort. The leader also continues to run the normal RAFT protocol - in the event
that a log entry remains unreplicated at a follower it sends it the log entry
for replication as usual.

The only change necessary for client assisted replication from the perspective
of RAFT is for follower replicas to be able to handle out of order arrival of
log entries. This is accomplished by an in memory cache of replicated log
entries. When an arriving entry completes a sequence of log entries starting at
the next desired index, the cache attempts to append the log entries to the
actual RAFT log. For each entry the appropriate RAFT response (accept/reject) is
sent back to the client that had sent the entry on behalf of the leader.

>>>>

In order to ease this bottleneck, Cyclone allows clients to assist in
replication. To do so with minimum changes to the RAFT protocol, the leader
continues to replicate entries to clients on a timeout, with the clients
providing best effort replication. The flow for assisted replication is
as follows. For any RPC call added to the log, the leader sends back a message
to the client with a list of replicas. The client then sends the RPC call to
follower replicas on behalf of the leader. It then then waits for a timeout,
during which it gathers responses from clients and finally sends back a message
to the leader with the list of replicas that have responded within the
timeout. Client assisted replication is therefore best effort. The leader also
continues to run the normal RAFT protocol - in the event
that a log entry remains unreplicated at a follower it sends it the log entry
for replication as usual.

The only change necessary for client assisted replication from the perspective
of RAFT is for follower replicas to be able to handle out of order arrival of
log entries. This is accomplished by an in memory cache of replicated log
entries - a structure we dub the log cache. When an arriving entry completes a
sequence of log entries starting at the next desired index, the cache attempts
to append the log entries to the actual RAFT log. For each entry the appropriate
RAFT response (accept/reject and the next desired log entry in the case of
accept) is sent back to the client that had sent the entry on behalf of the
leader. Pseudocode for the log cache is shown in Figure~\ref{fig:logcache}.

\begin{figure}
{ \scriptsize
\begin{verbatim}
on ReceiveEntry(term, entry) {
  if (term < logcache_term) {
    drop entry
  }
  else if(term > logcache_term){
    drop all entries from logcache
    logcache_term = term
  }
  while(entry = canAdd(raftLog, logcache)) {
    AddEntry(raftLog, entry);
    RemoveEntry(logcache, entry)
    Send response to client for entry;
  }
}
\end{verbatim}
}
\caption{Logcache}
\label{fig:logcache}
\end{figure}

Any client assists in the replication of exactly one entry at a time, dropping
the previous entry is it receives a new entry from the leader to replicate. The
client needs to therefore uniquely identify the entry for which a response is
received from a replica. In RAFT, entries being replicated are uniqely
identified by the term of the leader, the term of the entry and the index of the
previous entry. These pieces of identifying information are carried in
replication messages to allow RAFT to tolerate out of order and unreliable
messages on the network. The client uses these parameters to filter out all
reponses from replicas other than the one for the entry it is attempting to
replicate.

Finally, we need to show that the logcache does not lead to a departure from the
RAFT protocol. This follows from the fact that in RAFT there is exactly one
leader for any given term number (stemming from the safety properties of the
election protocol). Further, the leader always extens its log, it never rolls
back log entries - entries always flow from leader to followers in RAFT (unlike
other protocols such as Paxos). Therefore any log entry added to the logcache is
always present at the leader in that term. We ensure stale entries do not
survive leadership changes by flushing the cache on a change in term. Hence any
entry appended to the actual RAFT log at the follower is one that is present at
the leader - the protocol therefore behaves no differently than if the leader
had actually transmitted the log entry directly to the follower replica.
